---
title: "Assignment 1 - Network Analysis"
author: "Balint Bojko"
date: '2018 november 2 '
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


<style>
body {
text-align: justify}
</style>

##Question 1

###a,
The following code gives NA, because NA values in included in the computation process.

```{r}
a <- c(8, NA, 20, 5, 7, 9)
mean(a)
```

We need to drop the NA, by including na.rm in the mean function.

```{r}
mean(a,na.rm = T)
```

###b,
Now, remove the NA from the vector instead and compute the mean.
```{r}
a <- a[-2]
mean(a)
```

##Question 2
###a,
```{r}
matrix_Q2 <- matrix(c(0,1,1,1,rep(0,12)),4,4,byrow = T)
```

###b,
```{r}
library(qgraph)
qgraph(matrix_Q2)
```

##Question 3
###a,
```{r}
library(psych)
# Load bfi dataset:
data(bfi)
# Look at data:
View(bfi)
```
More information about the data set can be found by running '?bfi' from the console.

###b,
Number of men
```{r}
sum(bfi$gender==1)
```
Number of women
```{r}
sum(bfi$gender==2)
```

###c,
Mean age of subjects
```{r}
mean(bfi$age)
```

###d,
Found information about EFA function in the psych package description ('?psych') / ('?psych::fa')

Creating new dataset
```{r}
data_Q3 <- bfi[,1:25]
fa_Q3 <- fa(data_Q3, rotate = 'promax')
```
###e,
Creating plot with loadings
```{r}
qgraph.loadings(unclass(fa_Q3$loadings))
```


##Question 4
###a,
Setting the working directory from the menu: Session/Set Working Directory/To Source File Location

###b,
Loading SPSS file
```{r}
library(foreign)
data_Q4 <- read.spss('job_performance.sav',to.data.frame = T)
```

###c,
```{r}
cor(data_Q4[,-1])
```

###d,
Fitting linear regression with job performance as the outcome variable, and IQ, motivation and social support as predictors.
```{r}
fit <- lm(perf ~ iq + mot + soc, data = data_Q4)
```

###e,
Investigating object fit
```{r}
summary(fit)
```

All three predictors are significant with and without adjusting for multiple testing (Bonferroni). Quick question: do you need to adjust for multiple testing in this case?


##Question 5
###a,
```{r}
data_Q5 <- read.spss('PTSDdata.sav',to.data.frame = T)
data_Q5 <- data_Q5[,10:30]
colnames(data_Q5) <- c('Lifetime Trauma',
                       'Intrusive thoughts',
                       'Nightmares',
                       'Flashbacks',
                       'Emotional cue reactivity',
                       'Psychological cue reactivity',
                       'Avoidance of thoughts',
                       'Avoidance of reminders',
                       'Trauma-related amnesia',
                       'Negative beliefs',
                       'Blame of self or others',
                       'Negative trauma-related emotions',
                       'Loss of interest',
                       'Detachment',
                       'Restricted affect',
                       'Irritability / anger',
                       'Self-destructive / reckless behaviour',
                       'Hypervigilance',
                       'Exaggerated startle response',
                       'Difficulty concentrating',
                       'Sleep disturbance')
head(data_Q5)
```
###c,
```{r}
round(cor(data_Q5[,-1]),2)
```


###d,
```{r}
sum(data_Q5$Nightmares)
sum(data_Q5$`Sleep disturbance`)
```

###e,
```{r}
cor.test(data_Q5$`Lifetime Trauma`,data_Q5$Nightmares)
```
The correlation is significant, however, can be considered 'weak' only. Cor = 0.19, p = 0.005

###f,
```{r}
library(ppcor)
pcor.test(data_Q5$`Lifetime Trauma`,data_Q5$Nightmares,data_Q5$`Negative trauma-related emotions`)
```
The p-value is still significant at alpha = 0.05, however, it is higher with the control variable included. Cor = 0.164, p = 0.015. This happened because some of the shared variance between the two variables can be explained by the control variable.

##Bonus question
```{r, message=F, error=F,warning=F}
data_bonus <- read.csv('studentdata.csv',sep=';')

par(mfrow = c(1,1), mar = c(5,5,10,5))

barplot(prop.table(table(data_bonus$What.is.your.gender.)),  col=rgb(0,0,1,1/4),
        main = 'Gender')

barplot(prop.table(table(data_bonus$I.am.a.student.in.the.)),  col=rgb(0,0,1,1/4),
        main = 'Study programme')

barplot(prop.table(table(data_bonus$How.familiar.are.you.with.R.)),  col=rgb(0,0,1,1/4),
        main = 'Familiarity with R')

barplot(prop.table(table(data_bonus$What.is..are..your.specialization.s..)),  col=rgb(0,0,1,1/4),
        main = 'Specialization')

barplot(prop.table(table(data_bonus$Where.are.you.from.)),  col=rgb(0,0,1,1/4),
        main = 'Home country')

```





